### 2021-03-02 paper/letter.tex

\documentclass{letter}
\usepackage{hyperref}
\signature{Soledad Villar}
\address{Department of Mathematics and Statistics \\ Mathematical Institute for Data Science \\ Johns Hopkins University}
\begin{document}

\begin{letter}{Foundations of Data Science Journal}
\opening{Dear Editors:}

Please consider our manuscript entitled ``Fitting very flexible models: Linear regression with large numbers of parameters" for publication in your journal.

Our note is intended to be an expository review article directed to scientists and data scientists. We address fundamental questions about over-parameterized linear regression and their connections to kernel regression, which are workhorses of contemporary data science. In particular, we emphasize the importance of regularization through a compelling example, we discuss cross-validation, and implementation considerations. We address active research in machine learning, such as the double descent phenomenon.

\closing{Yours Faithfully,}

%\encl{Copyright permission form}

\end{letter}
\end{document}